A Standard for the Scholarly Citation of Archaeological Data

### 2017-10-06 analysis/paper/paper.Rmd
fix some typos spotted by Stefano Costa <steko@iosa.it>

output: 
  papaja::apa6_pdf:
   latex_engine: xelatex

@@

The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications. They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old. Vines et al. found that odds of a data set referred to in a publication being available fell by 17% per year. The found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.

@@

Our sample of authors that we wrote email to focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during 2009 and 2015. We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission. The data included with this paper is anonymised so the specific authors we contacted are not identifiable.

@@

Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`. We focused on stone artefact experiments because the all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission. The data included with this paper is anonymised so the specific authors we contacted are not identifiable.

We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and received `r emails_back` replies, resulting in a `r email_response_rate`% response rate. We received `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%. Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study. We found no clear relationship of data sharing and date of publication. We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication. Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation. We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised for ready for distribution. We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.

@@

This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting. Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands. BM was supported by an Australian Research Council Future Fellowship (FT140100101).

>>>>

output: 
  papaja::apa6_word 
   # :
   # latex_engine: xelatex

@@

The absence of a standardised and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications. They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old. Vines et al. found that odds of a data set referred to in a publication being available fell by 17% per year. They found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.

@@

@@

Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`. We focused on stone artefact experiments because all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission. The data included with this paper is anonymised so the specific authors we contacted are not identifiable.

We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and received `r emails_back` replies, resulting in a `r email_response_rate`% response rate. We received `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%. Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study. We found no clear relationship of data sharing and date of publication. We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication. Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation. We interpret these to mean that the author does not have time to prepare their data for sharing, perhaps because the data are not already in a well-organised form ready for distribution. We found no effect on the author's student status at time of publication on data sharing behaviours, with roughly equal proportions of students and non-students sharing and not sharing.

@@

This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting. Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands, and Stefano Costa for many helpful suggestions. BM was supported by an Australian Research Council Future Fellowship (FT140100101).
